---
title: "Deforestation"
author: "Hollie Rawlings"
date: "11/05/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidytuesdayR")
library(shiny)
library(tidyverse)
library(plotly)
library(janitor)
```

### Data call###
```{r}
#tuesdata <- tidytuesdayR::tt_load('2021-04-06')

#forest_data <- tuesdata$forest
#forest_area_data <-tuesdata$forest_area
#brazil_forest_loss <-tuesdata$brazil_loss
soybean_farming <-tuesdata$soybean_use
#vegtable_oil <-tuesdata$vegetable_oil
```

###Ideas###
1. A map of forest loss by country filterable by year (I have never done a map in R before)
2. Tree map (because we like those) of soybean use filterable by year and faceted by continent (means i will need to join in a continent list)
3. Forecast of future rainforest loss in Brazil for the next 10 years. 

### New ideas (soy themed)
1. Heatmap of soy production worldwide toggled by deforestation worldwide
2. Line chart over the years 
3. 

###Source###
paper is here: https://ourworldindata.org/forests-and-deforestation






### Tree map showing the changes in soybean use by continet over time
***Background and aims***
1. How has soy production changed over time 
2. However due to recent links with global deforestation, soy has been marked as a product to avoid. How does soy bean farming link to deforestation? 
3. Should I stop eating tofu to save trees? or is the real picture more complicated. 


categories are direct human consumption, direct animal feed and processed (animal feed, bio fuel,industrial use and vegetable oil)


1. I don't like how they have grouped continents in this code so I am joining my own list in
source is here: https://pkgstore.datahub.io/JohnSnowLabs/country-and-continent-codes-list/country-and-continent-codes-list-csv_csv/data/b7876b7f496677669644f3d1069d3121/country-and-continent-codes-list-csv_csv.csv

2. Next I look for a primary key iD to join the two on, in this case its the 3 letter code, I do an inner join so that all the things I don't want in soybean_farming are not included. Then Ill select what we need. 

3. Some countries exist in two continents Ill create a list of allocations so I can see where it is going wrong. 
4. Then I mutate the continent with case_when to give it one allocation
```{r}
continent_list <-read.csv("Continent_list.csv")

soybean_farming <- soybean_farming %>% 
                                      inner_join(continent_list,by=c("code"="Three_Letter_Country_Code"))%>%
                                      select(entity,code,year,human_food,animal_feed,processed,`ï..Continent_Name`)%>%
                                      rename(Continent=`ï..Continent_Name`)

continent_allocation <- soybean_farming %>% select(entity,Continent)%>%
                                                                        unique()
## These countries are 	Armenia,Azerbaijan,Cyprus,Georgia,Kazakhstan,Russia,Turkey 
## There are not many, so I can use a case when to recode them. 
## I'm no geography expert so I may not be right, If this were go go in a paper I would have to include a footnote about these allocations

soybean_farming <- soybean_farming %>% 
                                      mutate(Continent = case_when(
                                        entity == "Armenia" ~ "Europe",
                                        entity == "Azerbaijan" ~ "Asia", 
                                        entity == "Cyprus" ~ "Europe", 
                                        entity == "Georgia" ~ "Asia", 
                                        entity == "Kazakhstan" ~ "Asia", 
                                        entity == "Russia" ~ "Europe", 
                                        entity == "Turkey" ~ "Europe", 
                                        TRUE ~ Continent
                                        ))%>%
                                        unique()



```

## Data processing for suburst chart 
1. As a sample to just get myself going I have limited the year to 2013
2. Eventually year will go in as a slider or dropdown in the shiny app
3. I remove everything that we don't need in the data file and re-order it specifically for the data wrangling function
```{r}
soybean_farming <- soybean_farming %>%
                                        filter(year == "2013")%>%
                                        select(-year,-code)%>%
                                        pivot_longer(cols =c(human_food,animal_feed,processed))%>%
                                        select(Continent,entity,name,value)
                                      
```

## Function for building enginering data into the sunburst plot (its a real pain to get data in this format, ill keep note of this for the future) 
```{r}
as.sunburstDF <- function(DF, valueCol = NULL){
  require(data.table)
  
  colNamesDF <- names(DF)
  
  if(is.data.table(DF)){
    DT <- copy(DF)
  } else {
    DT <- data.table(DF, stringsAsFactors = FALSE)
  }
  
  DT[, root := "Total"]
  colNamesDT <- names(DT)
  
  if(is.null(valueCol)){
    setcolorder(DT, c("root", colNamesDF))
  } else {
    setnames(DT, valueCol, "values", skip_absent=TRUE)
    setcolorder(DT, c("root", setdiff(colNamesDF, valueCol), "values"))
  }
  
  hierarchyCols <- setdiff(colNamesDT, "values")
  hierarchyList <- list()
  
  for(i in seq_along(hierarchyCols)){
    currentCols <- colNamesDT[1:i]
    if(is.null(valueCol)){
      currentDT <- unique(DT[, ..currentCols][, values := .N, by = currentCols], by = currentCols)
    } else {
      currentDT <- DT[, lapply(.SD, sum, na.rm = TRUE), by=currentCols, .SDcols = "values"]
    }
    setnames(currentDT, length(currentCols), "labels")
    hierarchyList[[i]] <- currentDT
  }
  
  hierarchyDT <- rbindlist(hierarchyList, use.names = TRUE, fill = TRUE)
  
  parentCols <- setdiff(names(hierarchyDT), c("labels", "values", valueCol))
  hierarchyDT[, parents := apply(.SD, 1, function(x){fifelse(all(is.na(x)), yes = NA_character_, no = paste(x[!is.na(x)], sep = ":", collapse = " - "))}), .SDcols = parentCols]
  hierarchyDT[, ids := apply(.SD, 1, function(x){paste(x[!is.na(x)], collapse = " - ")}), .SDcols = c("parents", "labels")]
  hierarchyDT[, c(parentCols) := NULL]
  return(hierarchyDT)
}
sunburstDF <- as.sunburstDF(soybean_farming, valueCol = "value")
```
### Radial treemap 


```{r}
sunburst <-plot_ly(data = sunburstDF, ids = ~ids, labels= ~labels, parents = ~parents, values= ~values, type='sunburst',maxdepth=2 ,branchvalues = 'total')
sunburst

##Find a way to stop labels running into each other
```
